# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

We assume that the current directory contains the zipped data file 'activity.zip'. 

Unzip it and load it:
```{r echo=TRUE}
unzip('activity.zip')
df <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?

Aggregate the data to get the total of steps per day:

```{r echo=TRUE}
steps.per.day <- aggregate(df$steps, list(df$date), sum, na.rm=T)
names(steps.per.day) <- c("date", "num.steps")
```

Histogram of the steps taken per day

```{r echo=TRUE}
hist(steps.per.day$num.steps, breaks=10, main="Histogram: number of steps per day", xlab="number of steps per day")
```

Calculate mean ond median of the total number of steps per day:
```{r echo=TRUE}
mean(steps.per.day$num.steps, na.rm=T)
median(steps.per.day$num.steps, na.rm=T)
```


## What is the average daily activity pattern?

Aggregate the steps over the 5-minutes intervals, taking the mean over all days:

```{r echo=TRUE}
steps.per.interval <- aggregate(df$steps, list(df$interval), mean, na.rm=T)
names(steps.per.interval) <- c("interval", "mean.steps")
```

Plot the time series of the 5-minute interval and the average number of steps taken, averaged across all days.

```{r echo=TRUE}
plot(steps.per.interval$interval, steps.per.interval$mean.steps, t="l", main="Average steps per interval", ylab="Number of steps", xlab="Interval")
```

Show the 5-minutes intervall with the most steps (averaged over all days):

```{r echo=TRUE}
steps.per.interval[ steps.per.interval$mean.steps==max(steps.per.interval$mean.steps),]
```

## Imputing missing values

### Total number of missing values

The steps-field has 2304 missing values, whereas both the date and interval fields have no missing values:
```{r echo=TRUE}
sum(is.na(df$steps))
sum(is.na(df$date))
sum(is.na(df$interval))
```

### Replacing missing values

Copy the datasest to a new data.frame `df.complete` and replace missing values with the mean of the 5-minutes-interval averaged over all days.

```{r echo=TRUE}

df.complete <- df
for (i in 1:nrow(df)) {
  if ( is.na(df[i, 1]) ) {
    df.complete[i, 1] <- steps.per.interval[ steps.per.interval$interval==df[i, 3], ][2]
  }
}
```

### Histogram of the steps taken per day, based on replaced missing values

```{r echo=TRUE}
steps.per.day.complete <- aggregate(df.complete$steps, list(df.complete$date), sum, na.rm=T)
names(steps.per.day.complete) <- c("date", "num.steps")

hist(steps.per.day.complete$num.steps, breaks=10, main="Histogram: number of steps per day", xlab="number of steps per day")
```

### Mean and median of steps taken per day, based on replaced missing values

Calculate mean ond median of the total number of steps per day:
```{r echo=TRUE}
mean(steps.per.day.complete$num.steps, na.rm=T)
median(steps.per.day.complete$num.steps, na.rm=T)
```

#### Changes introduced by replacing missing values: 
Days without a single steps measurement (i.e. only NAs) show up as having 0 steps in the histogram.   
With missing values replaced by the mean of the interval over the days, such days with seemingly
0 steps disappear.    
With missing values, the mean is 9354 and the median 10395.    
Without missing values, both mean and median are 10766.

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r echo=TRUE}
weekend <- weekdays( as.Date(df.complete$date)) == "Saturday" | weekdays( as.Date(df.complete$date)) == "Sunday"
df.complete$weekend <- as.factor(weekend)
```

Number of steps per interval averaged over weekdays and over weekends:

```{r echo=TRUE}
df.weekday <- df.complete[ df.complete$weekend==FALSE, ]
df.weekend <- df.complete[ df.complete$weekend==TRUE, ]
weekday.steps <- aggregate(df.weekday$steps, list(df.weekday$interval), mean, na.rm=T)
weekend.steps <- aggregate(df.weekend$steps, list(df.weekend$interval), mean, na.rm=T)
steps.per.interval$weekdays <- weekday.steps$x
steps.per.interval$weekend  <- weekend.steps$x

par(mfrow=c(2,1))
plot(steps.per.interval$interval, steps.per.interval$weekend, t="l", main="weekend", ylab="Number of steps", xlab="Interval")
plot(steps.per.interval$interval, steps.per.interval$weekdays, t="l", main="weekdays", ylab="Number of steps", xlab="Interval")

```

### A simpler way to do the weekend/weekday plot:

```{r echo=TRUE}
spiw <- aggregate(df.complete$steps, list(df.complete$weekend, df.complete$interval), mean, na.rm=T)
library(lattice)
names(spiw) <- c("weekend", "interval", "mean.steps")
xyplot(spiw$mean.steps ~ spiw$interval | spiw$weekend)
```
spiw <- aggregate(df.complete$steps, list(df.complete$weekend, df.complete$interval), mean, na.rm=T)