# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

We assume that the current directory contains the zipped data file 'activity.zip'. 

Unzip it and load it, and discard rows :
```{r echo=TRUE}
unzip('activity.zip')
df <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?

Aggregate the data to get the total of steps per day:

```{r echo=TRUE}
num.steps.per.day <- aggregate(df$steps, list(df$date), sum, na.rm=T)
names(num.steps.per.day) <- c("date", "num.steps")
num.steps.per.day
```

Aggregate the data to get the mean number of steps per day:

```{r echo=TRUE}
mean.steps.per.day <- aggregate(df$steps, list(df$date), mean, na.rm=T)
names(mean.steps.per.day) <- c("date", "mean.steps")
mean.steps.per.day
```

Aggregate the data to get the median number of steps per day:

```{r echo=TRUE}
median.steps.per.day <- aggregate(df$steps, list(df$date), median, na.rm=T)
names(median.steps.per.day) <- c("date", "median.steps")
median.steps.per.day
```

### Histogram of the steps taken per day

```{r echo=TRUE}
hist(num.steps.per.day$num.steps, breaks=10, main="Histogram: number of steps per day", xlab="number of steps per day")
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
